A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-05-01 paper.tex
Fix remaining issues with bib file

In addition, scatterplots of residuals plotted against the fitted values
and each of the explanatory variables, can be used to scrutinize their
relationships. If there are any visually discoverable patterns, the
model is potentially inadequate or incorrectly specified. In general,
one looks for noticeable departures from the model such as non-linear
pattern or heteroskedasticity. A non-linear pattern would suggest that
the model needs to have some additional non-linear terms.
Heteroskedasticity suggests that the error is dependent on the
predictors, and hence violates the independence assumption. However,
correctly judging whether NO pattern exists in a residual plot is a
difficult task for humans. Humans will almost always will see a pattern
(see \citet{kahneman}), so the question that really needs answering is
whether any pattern perceived is consistent with randomness, purely
sampling variability, or noise. It is especially difficult to teach this
to new analysts and students \citep{loy2021bringing}. To address this,
statistical tests are available to check for non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.

@@

With large sample sizes, hypothesis tests may reject the null hypothesis
when there is only a small effect. (A good discussion can be found in
\citet{kirk1996}.) While such rejections may be statistically correct,
their sensitivity may render the results impractical. A key goal of
residual plot diagnostics is to identify potential issues that could
lead to incorrect conclusions or errors in subsequent analyses, but
minor defects in the model are unlikely to have a significant impact and
may be best disregarded for practical purposes. The experiment discussed
in this paper specifically addresses this tension between statistical
and practical statistics.

>>>>

In addition, scatterplots of residuals plotted against the fitted values
and each of the explanatory variables, can be used to scrutinize their
relationships. If there are any visually discoverable patterns, the
model is potentially inadequate or incorrectly specified. In general,
one looks for noticeable departures from the model such as non-linear
pattern or heteroskedasticity. A non-linear pattern would suggest that
the model needs to have some additional non-linear terms.
Heteroskedasticity suggests that the error is dependent on the
predictors, and hence violates the independence assumption. However,
correctly judging whether NO pattern exists in a residual plot is a
difficult task for humans. Humans will almost always will see a pattern
(see \citet{kahneman2011thinking}), so the question that really needs
answering is whether any pattern perceived is consistent with
randomness, purely sampling variability, or noise. It is especially
difficult to teach this to new analysts and students
\citep{loy2021bringing}. To address this, statistical tests are
available to check for non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.

@@

With large sample sizes, hypothesis tests may reject the null hypothesis
when there is only a small effect. (A good discussion can be found in
\citet{kirk1996practical}.) While such rejections may be statistically
correct, their sensitivity may render the results impractical. A key
goal of residual plot diagnostics is to identify potential issues that
could lead to incorrect conclusions or errors in subsequent analyses,
but minor defects in the model are unlikely to have a significant impact
and may be best disregarded for practical purposes. The experiment
discussed in this paper specifically addresses this tension between
statistical and practical statistics.
